BivData_Setup

### 2020-01-17 src/BivData_Setup.Rmd
working on the BivData setup

---
title: "BivData_Setup"
author: "April Peterson"
date: "January 17, 2020"
output: html_document
---

```{r setup, include=FALSE}

setwd("~./MLH1repo/")
source("src/CommonFunc_MLH1repo.R")

#BIVDATA
new.Curate.FULL <- read.csv("data/BivData/curation/FULL_Curated_BivData.csv")
#80919
#61084
#60241
#60830
#57548

new.Curate.FULL <- new.Curate.FULL[!(is.na(new.Curate.FULL$fileName) | new.Curate.FULL$fileName==""), ]
new.Curate.FULL$Obj.ID <- paste(new.Curate.FULL$fileName, new.Curate.FULL$boxNumber, sep = "_")
new.Curate.FULL <- add_mouse(new.Curate.FULL)

#make OG copy
new.Curate.FULL.org <- new.Curate.FULL

new.Curate.FULL.org <- add_category(new.Curate.FULL.org)

OG.cURATE.table.mouse <- table(new.Curate.FULL.org$mouse, new.Curate.FULL.org$SC.pass)
OG.cURATE.table.category <- table(new.Curate.FULL.org$category, new.Curate.FULL.org$SC.pass)

OG.cURATE.table.mouse <- add_strain(OG.cURATE.table.mouse)

#write the curated file
write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_17.1.20.csv", sep=",",
            row.names = TRUE)

write.table(OG.cURATE.table.category, "~./MLH1repo/data/CurrentCuratedDataset_cat_17.1.20.csv", sep=",",
            row.names = TRUE)

@@

@@

```{r CLEAN.UP, echo=FALSE, include=FALSE}

@@

# basic stats

@@

@@

@@

# analysis

@@

@@

- SC

>>>>

---
title: "BivData_Setup"
author: "April Peterson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
library(plyr)
library(lattice)
library(dplyr)
library(ggplot2)
library(reshape2)

setwd("~./MLH1repo/")
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

#BIVDATA MASTER SHEET
new.Curate.FULL <- read.csv("~./MLH1repo/data/BivData/curation/FULL_Curated_BivData.csv")
#80919
#61084
#60241
#60830
#57548

#Clean away all the empty rows
new.Curate.FULL <- new.Curate.FULL[!(is.na(new.Curate.FULL$fileName) | new.Curate.FULL$fileName==""), ]
new.Curate.FULL$Obj.ID <- paste(new.Curate.FULL$fileName, new.Curate.FULL$boxNumber, sep = "_")
colnames(new.Curate.FULL)
new.Curate.FULL <- add_mouse(new.Curate.FULL)

#make OG copy
new.Curate.FULL.org <- new.Curate.FULL
new.Curate.FULL.org <- add_category(new.Curate.FULL.org)

#write files for basic stats to csv files
OG.cURATE.table.mouse <- table(new.Curate.FULL.org$mouse, new.Curate.FULL.org$SC.pass)
#OG.cURATE.table.mouse <- add_strain(OG.cURATE.table.mouse)

OG.cURATE.table.category <- table(new.Curate.FULL.org$category, new.Curate.FULL.org$SC.pass)

#write the curated file
write.table(OG.cURATE.table.mouse, "~./MLH1repo/data/CurrentCuratedDataset_17.1.20.csv", sep=",",
            row.names = TRUE)
#
write.table(OG.cURATE.table.category, "~./MLH1repo/data/CurrentCuratedDataset_cat_17.1.20.csv", sep=",",
            row.names = TRUE)

#SC PASS
new.Curate.FULL <- new.Curate.FULL[new.Curate.FULL$SC.pass == '1',]#8933

@@

Goal of this doc is to make DFs that are ready for the analysis.

Main files from this plot:

  1. new.Curate.FULL: cleaned up version of the passing, curated Bivalent measures

  2. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)

  3. MELT_ABS: make melt version of foci positions for abs measuers

  4. MELT_PER: USE of the general foci position plots

@@

```{r CLEAN.UP.1, echo=FALSE, include=FALSE}

#check for duplicates
anyDuplicated(new.Curate.FULL$Obj.ID)# 3467

dup.curated <- new.Curate.FULL[duplicated(new.Curate.FULL$Obj.ID),] #784
dup.curated <- add_mouse(dup.curated)

table(dup.curated$mouse)

#which lines are duplicated
#lots of PWD_female duplicated -- probably how I put together the curated files

#fix these duplicated ones

#remove duplicated
Curated_BivData <- new.Curate.FULL %>% distinct()#from 8933 to 8151

#not surethe value of geting these lines
#new.Curate.FULL <- new.Curate.FULL[ !grepl("[?]", new.Curate.FULL$SC.pass) , ]
#new.Curate.FULL <- new.Curate.FULL[ !grepl("[?]", new.Curate.FULL$hand.foci.count) , ]

#new.Curate.FULL <- new.Curate.FULL.org[(new.Curate.FULL.org$SC.pass == 1 | new.Curate.FULL.org$SC.pass == 0),]
#new.Curate.FULL <- droplevels(new.Curate.FULL)

#new.Curate.FULL <- add_mouse(new.Curate.FULL)
#new.Curate.FULL <- add_category(new.Curate.FULL)

#pass.fail.table <- table(new.Curate.FULL$SC.pass, new.Curate.FULL$category)
#this outsput shows the relative ratios of pass to fail biv segmentation

Curated_BivData <- add_mouse(Curated_BivData)
Curated_BivData <- add_category(Curated_BivData)
Curated_BivData <- add_strain(Curated_BivData)
Curated_BivData <- add_sex(Curated_BivData)
Curated_BivData <- add_subsp(Curated_BivData)
```

```{r set.type, echo=FALSE}

##change factor cols to numbers (factor -> chr -> number)
Curated_BivData$centromere_ABS_Position <- as.character(Curated_BivData$centromere_ABS_Position)
Curated_BivData$centromere_PER_Position <- as.character(Curated_BivData$centromere_PER_Position)
Curated_BivData$chromosomeLength <- as.character(Curated_BivData$chromosomeLength)

Curated_BivData$hand.foci.count <- as.character(Curated_BivData$hand.foci.count)

Curated_BivData$Foci1 <- as.character(Curated_BivData$Foci1)
Curated_BivData$Foci2 <- as.character(Curated_BivData$Foci2)
Curated_BivData$Foci3 <- as.character(Curated_BivData$Foci3)

##
Curated_BivData$centromere_ABS_Position <- as.numeric(Curated_BivData$centromere_ABS_Position)
Curated_BivData$centromere_PER_Position <- as.numeric(Curated_BivData$centromere_PER_Position)
Curated_BivData$chromosomeLength <- as.numeric(Curated_BivData$chromosomeLength)

Curated_BivData$hand.foci.count <- as.numeric(Curated_BivData$hand.foci.count)

Curated_BivData$Foci1 <- as.numeric(Curated_BivData$Foci1)
Curated_BivData$Foci2 <- as.numeric(Curated_BivData$Foci2)
Curated_BivData$Foci3 <- as.numeric(Curated_BivData$Foci3)
#I think the number of sig figs and rounding changes slightly from the blocks above

#adding more feature calculations (IFD, ect)
Curated_BivData <- add_IFD(Curated_BivData)
Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength

#IFD.2 does mulitple IFDs (just absolute)
#Curated_BivData <- add_IFD.2(Curated_BivData)

Curated_BivData$PER_Foci_1 <- Curated_BivData$Foci1 / Curated_BivData$chromosomeLength
Curated_BivData$PER_Foci_2 <- Curated_BivData$Foci2 / Curated_BivData$chromosomeLength
Curated_BivData$PER_Foci_3 <- Curated_BivData$Foci3 / Curated_BivData$chromosomeLength

Curated_BivData$PER_Foci_1 <- as.numeric(Curated_BivData$PER_Foci_1)
Curated_BivData$PER_Foci_2 <- as.numeric(Curated_BivData$PER_Foci_2)
Curated_BivData$PER_Foci_3 <- as.numeric(Curated_BivData$PER_Foci_3)

Curated_BivData$dis.cent <- Curated_BivData$Foci1 - Curated_BivData$centromere_ABS_Position
Curated_BivData$dis.cent.PER  <- Curated_BivData$dis.cent / Curated_BivData$chromosomeLength

str(Curated_BivData)
```

IFD functions

  - IFD: currently I think just does 2CO bivalents

  - IFD.2: for mulitple IFDs (3CO ect.)

```{r CLEAN.3}

#report some of these things in the text, for the report

#remove data which has paint marks
P_fileNames <- Curated_BivData$fileName[(grep('p_rev', Curated_BivData$fileName))]
length(P_fileNames)

#what is this 
#prev_num = length(P_fileNames)
#fullBivData <- fullBivData[!(fullBivData$fileName %in% P_fileNames), ]

#remove anything with 'DUPLICATE'
DUP_list <- Curated_BivData$fileName[(grep('DUP', Curated_BivData$fileName))]
DUP_num = length(DUP_list)#0

#fullBivData <- fullBivData[!(fullBivData$fileName %in% DUP_list), ]

#remove images that would be doubles ... 12, 12.1, 12.2
#this isn't completely effective

dup_list= c()
for(i in Curated_BivData$fileName){  #change the file name header,  
  # print(i)
  ifelse(grepl("\\.[0-9]_rev", i),
         (dup_list = c(dup_list, i)), NA)
}
MultiCells = c()
for(j in dup_list){
  stringLIST = strsplit(j, split="\\.")[[1]]
  new_string = paste(stringLIST[[1]][1], "rev", sep="_")
  MultiCells = c(MultiCells, new_string)
}

MultiCells = unique(MultiCells)
multi_cell_num = length(MultiCells)#118

#test this part of the function
#I think most of these cell lines have been left out of the final DF
fullBivData <- fullBivData[!(fullBivData$fileName %in% MultiCells), ]
#save these number measures to report

@@

remove DUP, 'p', remove mulit cells (10, 10.1, 10.2)

what colnames do all the OFD functions need?

Discribe which images / Bivalents are removed

- Add F4 column!

# Basic stats

- pass rate

- number per mice?

Plots

1. Histograms of N.bivs per cell, facet category

2. Number of images per mouse, number of biavlents per mouse (not sure what kinda of plot for this)

3.

Is there a way to check my assumptions for these measues?

What are the sources of error fot these data?

```{r cell.table1, echo=FALSE}

cell.level.BivData_table <- ddply(.data=Curated_BivData, 
                       .(fileName),
                       summarize, 
                       boxs.IDd = max(boxNumber),
                       nboxes_passed = length(boxNumber),

                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
                       CO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
                       CO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
                       CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
                       CO3 = sum(hand.foci.count == 3, na.rm=TRUE),

                       ncells = length(unique(fileName)),
                       nbivs = length(unique(Obj.ID)),

#
 #                      mean.SC_length = mean(chromosomeLength),
                       var.SC_length = var(chromosomeLength),
                       sd.SC_length   = round(sd(chromosomeLength), 3),
                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3)
)   
                       #add the centromere and ifd here?
                       #mean.dist_cent = mean(dis.cent, na.rm=TRUE),
#                       var.dist_cent = var(dis.cent, na.rm=TRUE),
#                       sd.dist_cent   = round(sd(dis.cent, na.rm=TRUE), 3),
#                       se.dist_cent   = round(sd.dist_cent / sqrt(nbivs), 3),

#                       mean.siscoten = mean(SisCoTen, na.rm=TRUE),
#                       var.siscoten = var(SisCoTen, na.rm=TRUE),
#                       sd.siscoten   = round(sd(SisCoTen, na.rm=TRUE), 3),
 #                      se.siscoten   = round(sd.siscoten / sqrt(nbivs), 3),

#                       mean.IFD = mean(IFD1, na.rm=TRUE),#IFD1 is chr
#                       var.IFD = var(IFD1, na.rm=TRUE),
#                       sd.IFD   = round(sd(IFD1, na.rm=TRUE), 3),
#                       se.IFD   = round(sd.IFD / sqrt(nbivs), 3),

 #                      mean.IFDper = mean(IFD1_PER, na.rm=TRUE),#IFD1 is chr
#                       var.IFDper = var(IFD1_PER, na.rm=TRUE),
#                       sd.IFDper   = round(sd(IFD1_PER, na.rm=TRUE), 3),
#                       se.IFDper   = round(sd.IFDper / sqrt(nbivs), 3)

cell.level.BivData_table <- add_mouse(cell.level.BivData_table)
cell.level.BivData_table <- add_strain(cell.level.BivData_table)
cell.level.BivData_table <- add_category(cell.level.BivData_table)

@@

decide on the colnames that I want for the final DF

fileName, Object_ID, box.number, centromere_ABS, centromere_PER
chromosome length, SC.pass, Foci.pass, hand.foci.count, 
F1, F2, F3, F4, F1_per, F2_per, F3_per, F4_per, IFD
mouse, strain, category, subsp,

merge:
cell.MLH1.count, cell.quality, mouse.age, 
cell.total_skel, cell.total_binary

@@

```{r melt.table}

#paste in this code

coco.names <- colnames(Curated_BivData)

#"fileName" ,  "chromosomeLength",        "centromere_ABS_Position", "centromere_PER_Position",
#"boxNumber",               "Obj.ID",                  "SC.pass",                 "foci.pass",              
#"curated.1pass.0fail",     "curation.notes",          "hand.foci.count",        "Foci1",                  
#"Foci2",       "Foci3",                "notes",                 "mouse",                  
#"category",                "strain",                  "sex",                    "subsp",                  
#"IFD1_ABS",                "IFD1_PER",

#all of the cols to use
MELT_Curated_BivData <- melt(Curated_BivData[,c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",  "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse",
              #leave out the Foci
                "Foci1","Foci2","Foci3")],

            id=c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",
                    "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse"))

#dm1$value is the number for the foci position
#clean up dm1 (there are a bunch of rows without useful information)

MELT_Curated_BivData <- MELT_Curated_BivData[(!is.na(MELT_Curated_BivData$value)),]

#MELT, -- 15
colnames(MELT_Curated_BivData) <- c(coco.names[c(1:11, 15:16)], "Foci_Type", "Position") #too many foci
#I'm missing a number of cols -- not

```{r MELT.PER}

#run the melting for the PER foci position

colnames(Curated_BivData)

MELT_Curated_BivData.PER <- melt(Curated_BivData[,c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",  "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse",
              #leave out the Foci
                "PER_Foci_1","PER_Foci_2","PER_Foci_3")],

            id=c("fileName","chromosomeLength","centromere_ABS_Position","centromere_PER_Position",
                    "boxNumber", "Obj.ID", "SC.pass","foci.pass",              
              "curated.1pass.0fail","curation.notes", "hand.foci.count","notes", "mouse"))

MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[(!is.na(MELT_Curated_BivData.PER$value)),]

#MELT, -- 15
colnames(MELT_Curated_BivData.PER) <- c(coco.names[c(1:11, 15:16)], "Foci_Type", "PER_Position")

#figure out the different hand.foci.count

```{r merge.melt}
#merge the 2 melted DF
FULL_MELT_Curated_BivData  <- cbind( MELT_Curated_BivData, MELT_Curated_BivData.PER[15])

colnames(FULL_MELT_Curated_BivData) <- c(FULL_MELT_Curated_BivData[1:14], "Pos_ABS", "Pos_PER")

#
MELT_Curated_BivData.PER <- add_strain(MELT_Curated_BivData.PER)
MELT_Curated_BivData.PER <- add_sex(MELT_Curated_BivData.PER)

```{r general.foci.pos, echo=FALSE}

#make the normalized foci density plots
MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[MELT_Curated_BivData.PER$hand.foci.count < 5, ]
MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[!MELT_Curated_BivData.PER$hand.foci.count == 0, ]

MELT_Curated_BivData.PER <- MELT_Curated_BivData.PER[MELT_Curated_BivData.PER$PER_Position < 1.1, ]

#remove 0 and NA
#remove 4foci count?

#check for values over 1

density <- ggplot(MELT_Curated_BivData.PER, aes(x=PER_Position, fill=sex) ) + geom_density(alpha=.4) + facet_wrap(~hand.foci.count)+ggtitle("Normalized Foci Positions")

density
#table(the4$mouse)
#something wrong with the 4 bivalents
#many of the lines just have 3 lines

@@

```{r save.file}

#save Rdata
save.image("~./MLH1repo/data/CleanBivData.RData")
#save.image("data/BivData_4_PCA.RData")

# Number of whole cells

- list the cells with all bivs measued or which are close to full.

# Analysis

@@

Use the Bivalent_Analysis.Rmd for analysis pieces
BivPattern.Rmd started to analyze this but, not a very clear setup

@@

- SC lengths, by CO number

-ge
